### 2021-10-13 manuscript/draft-supplement.Rmd

---
title: "Comparative vaccine effectiveness of ChAdOx1 versus BNT162b2 in Health and Social Care workers in England: Supplementary materials"
output:
  pdf_document:
    toc: no
  word_document: default
  bookdown::html_document2:
    number_sections: no
    toc: no
  html_document:
    keep_md: yes
    self_contained: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
bibliography: references.bib
---

```{r setup, include=FALSE}

library('tidyverse')
library('here')
library('glue')
library('lubridate')
library('gt')
library('patchwork')
library('scales')
source(here("analysis", "lib", "utility_functions.R"))

#output_dir <- here("output")
output_dir <- here("released_outputs", "output")

fs::dir_create(output_dir, "report", "figures")

knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(fs::path(output_dir, "report", "figures"), "/"),
  fig.pos = 'H' #to stop figures floating around when rendering to pdf
)

gbl_vars <- jsonlite::read_json(
  path=here("output", "data", "metadata_study-dates.json")
)

flowchart <- read_csv(fs::path(output_dir, "data", "flowchart.csv"))

metadata_outcomes <- read_rds(here("output", "data", "metadata_outcomes.rds"))

list_formula <- read_rds(here("output", "data", "metadata_formulas.rds"))
list2env(list_formula, globalenv())
lastfupday <- lastfupday20
```

```{r characteristics, echo=FALSE, warning=FALSE, message=FALSE}
baseline <- read_csv(fs::path(output_dir, "report", "objects", "baseline.csv"))
baseline_az <- read_csv(fs::path(output_dir, "report", "objects", "baseline_az.csv"))
baseline_pfizer <- read_csv(fs::path(output_dir, "report", "objects", "baseline_pfizer.csv"))
```

```{r, effects, echo=FALSE, message=FALSE, warning=FALSE}

  outcomes <- 
    c(
      "postest",
      "covidemergency",
      "covidadmitted"
    ) %>% 
    set_names(., .)

  cmlinc_ns <- read_csv(fs::path(output_dir, "report", "objects", "cmlinc_ns.csv")) %>%
  mutate(
    outcome_descr = fct_inorder(outcome_descr),
    model_name = fct_inorder(str_wrap(model_name, 15))
  )
  cmlinc_pw <- read_csv(fs::path(output_dir, "report", "objects", "cmlinc_pw.csv")) %>%
  mutate(
    outcome_descr = fct_inorder(outcome_descr),
    model_name = fct_inorder(str_wrap(model_name, 15))
  )

  riskdiff_ns <- read_csv(fs::path(output_dir, "report", "objects", "riskdiff_ns.csv")) %>%
  mutate(
    outcome_descr = fct_inorder(outcome_descr),
    model_name = fct_inorder(str_wrap(model_name, 15))
  )
  riskdiff_pw <- read_csv(fs::path(output_dir, "report", "objects", "riskdiff_pw.csv")) %>%
  mutate(
    outcome_descr = fct_inorder(outcome_descr),
    model_name = fct_inorder(str_wrap(model_name, 15))
  )

  hazard_ns <- read_csv(fs::path(output_dir, "report", "objects", "hazard_ns.csv")) %>%
  mutate(
    outcome_descr = fct_inorder(outcome_descr),
    model_name = fct_inorder(str_wrap(model_name, 15))
  )
  hazard_pw <- read_csv(fs::path(output_dir, "report", "objects", "hazard_pw.csv")) %>%
  mutate(
    outcome_descr = fct_inorder(outcome_descr),
    model_name = fct_inorder(str_wrap(model_name, 15))
  )

  plot_cmlinc <- function(model){
    modell <- model

    cmlinc_ns %>%
      filter(model == modell) %>%
      ggplot()+
      geom_step(aes(x=tstop, y=(1-survival)*1000, group=vax1_az_descr, colour=vax1_az_descr))+
      geom_rect(aes(xmin=tstop, xmax=lead_tstop, ymin=(1-survival.ll)*1000, ymax=(1-survival.ul)*1000, group=vax1_az_descr, fill=vax1_az_descr), alpha=0.1)+
      geom_hline(aes(yintercept=0), colour='black')+
      facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
      scale_x_continuous(
        breaks = seq(0,7*52,by=28),
        expand = expansion(0)
      )+
      scale_y_continuous(
        expand = expansion(0),
        #limits = c(0L,NA)
      )+
      scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
      scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
      labs(
        x=NULL,
        y="Marginalised\ncumul. incidence\nper 1000",
        colour=NULL
      )+
      theme_minimal(base_size = 9)+
      theme(
        legend.position=c(0.95,0.95),
        legend.justification = c(1,1),
        legend.direction = "vertical",
        axis.text.x.top=element_text(hjust=0),

        panel.border = element_blank(),
        axis.line.y = element_line(colour = "black"),
        axis.line.x.top = element_line(colour = "black"),

        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background = element_blank(),

        panel.spacing = unit(0.8, "lines"),

      )+
      NULL
  }

   plot_riskdiff <- function(model){

    modell <- model
    riskdiff_ns %>%
      filter(model==modell) %>%
      ggplot()+
      geom_hline(aes(yintercept=0), colour='black')+
      geom_step(aes(x=tstop, y=diff*1000), colour='orchid4')+
      geom_rect(aes(xmin=tstop, xmax=lead_tstop, ymin=diff.ll*1000, ymax=diff.ul*1000), alpha=0.1, colour="transparent", fill='orchid4')+
      facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
      scale_x_continuous(
        limits = c(0, NA),
        breaks = seq(0,7*52,by=28),
        expand = expansion(0)
      )+
      scale_y_continuous(
        expand = expansion(0),
        sec.axis = dup_axis(name="Lower risk after \n <- BNT162b2 / ChAdOx1 ->", breaks = NULL)
      )+
      coord_cartesian(ylim= c(-3, 3))+
      scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
      scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
      labs(
        x=NULL,
        y="Difference\nin cumul. incidence\nper 1000",
        colour=NULL
      )+
      theme_minimal(base_size = 9)+
      theme(
        axis.line.y.left = element_line(colour = "black"),

        #panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),

        panel.spacing = unit(0.8, "lines"),

        plot.title = element_text(hjust = 0),
        plot.title.position = "plot",
        plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic")

       ) +
     NULL
  }

  plot_hazard <- function(model){
    modell <- model

    hazard_ns %>% 
      filter(model == modell) %>%
      ggplot()+
      geom_hline(aes(yintercept=1), colour='black')+
      geom_step(aes(x=tstop, y=hr), colour='orchid4', direction="vh")+
      geom_rect(aes(xmin=lag_tstop, xmax=tstop, ymin=hr.ll, ymax=hr.ul), alpha=0.1, colour="transparent", fill='orchid4')+
      facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
      scale_x_continuous(
        limits = c(0, NA),
        breaks = seq(0,7*52,by=28),
        expand = expansion(0)
      )+
      scale_y_log10(
        breaks=c(0.1, 0.2, 0.5, 1, 2, 5, 10),
        sec.axis = dup_axis(name="Lower hazard after \n <- BNT162b2 / ChAdOx1 ->", breaks = NULL)
      )+
      coord_cartesian(ylim= c(0.1, 10))+
      labs(
        x = "Days since first dose",
        y = "Hazard ratio"
      )+
      theme_minimal(base_size = 9)+
      theme(
        axis.line.y.left = element_line(colour = "black"),

        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),

        panel.spacing = unit(0.8, "lines"),

        plot.title = element_text(hjust = 0),
        plot.title.position = "plot",
        plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic")

       ) +
     NULL
  }

```

\newpage

### Figure S1a: Comparative effectiveness adjusting for region-specific calendar-time

```{r comp1, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', fig.height=7, out.extra=''}
patchwork::wrap_plots(plot_cmlinc(1), plot_riskdiff(1), plot_hazard(1), ncol=1)
```

\newpage

### Figure S1b: Comparative effectiveness adjusting for region-specific calendar-time and demographic characteristics

```{r comp2, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', fig.height=7, out.extra=''}
patchwork::wrap_plots(plot_cmlinc(2), plot_riskdiff(2), plot_hazard(2), ncol=1)
```

\newpage

### Figure S2: Fully adjusted comparative effectiveness using piecewise-constant estimates of the hazard

```{r plot_effects_pw, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', fig.height=7, results='asis', out.extra=''}

  plot_cmlinc_pw <- 
    cmlinc_pw %>%
    filter(model==3) %>%
    ggplot()+
    geom_step(aes(x=tstop, y=(1-survival)*1000, group=vax1_az_descr, colour=vax1_az_descr))+
    geom_rect(aes(xmin=tstop, xmax=lead_tstop, ymin=(1-survival.ll)*1000, ymax=(1-survival.ul)*1000, group=vax1_az_descr, fill=vax1_az_descr), alpha=0.1)+
    geom_hline(aes(yintercept=0), colour='black')+
    facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
    scale_x_continuous(
      breaks = seq(0,7*52,by=28),
      expand = expansion(0)
    )+
    scale_y_continuous(
      expand = expansion(0),
      #limits = c(0L,NA)
    )+
    scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
    labs(
      x=NULL,
      y="Marginalised\ncumul. incidence\nper 1000",
      colour=NULL
    )+
    theme_minimal(base_size = 9)+
    theme(
      legend.position=c(0.05,0.1),
      legend.justification = c(0,0),
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),

      panel.spacing = unit(0.8, "lines"),

    )+
    NULL

  ## difference in cumulative incidence

  plot_riskdiff_pw <-
    riskdiff_pw %>%
    filter(model==3) %>%
    ggplot()+
    geom_hline(aes(yintercept=0), colour='black')+
    geom_step(aes(x=tstop, y=diff*1000), colour='orchid4')+
    geom_rect(aes(xmin=tstop, xmax=lead_tstop, ymin=diff.ll*1000, ymax=diff.ul*1000), alpha=0.1, colour="transparent", fill='orchid4')+
    facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
    scale_x_continuous(
      limits = c(0, NA),
      breaks = seq(0,7*52,by=28),
      expand = expansion(0)
    )+
    scale_y_continuous(
      expand = expansion(0),
      sec.axis = dup_axis(name="Higher incidence after \n <- ChAdOx1 / BNT162b2 ->", breaks = NULL)
    )+
    coord_cartesian(ylim= c(-3, 3))+
    scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
    labs(
      x=NULL,
      y="Difference\nin cumul. incidence\nper 1000",
      colour=NULL
    )+
    theme_minimal(base_size = 9)+
    theme(
      axis.line.y.left = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),

      panel.spacing = unit(0.8, "lines"),

      plot.title = element_text(hjust = 0),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.caption = element_text(hjust = 0, face= "italic")

     ) +
   NULL

  ## hazard ratio
  plot_hazard_pw <-
    hazard_pw %>%
    filter(model==3) %>%
    ggplot()+
    geom_hline(aes(yintercept=1), colour='black')+
    geom_step(aes(x=tstop, y=hr), colour='orchid4', direction="vh")+
    geom_rect(aes(xmin=lag_tstop, xmax=tstop, ymin=hr.ll, ymax=hr.ul), alpha=0.1, colour="transparent", fill='orchid4')+
    facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
    scale_x_continuous(
      limits = c(0, NA),
      breaks = seq(0,7*52,by=28),
      expand = expansion(0)
    )+
    scale_y_log10(
      breaks=c(0.1, 0.2, 0.5, 1, 2, 5, 10),
      sec.axis = dup_axis(name="Higher hazard after \n <- ChAdOx1 / BNT162b2 ->", breaks = NULL)
    )+
    coord_cartesian(ylim= c(0.1, 10))+
    labs(
      x = "Days since first dose",
      y = "Hazard ratio"
    )+
    theme_minimal(base_size = 9)+
    theme(
      axis.line.y.left = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),

      panel.spacing = unit(0.8, "lines"),

      plot.title = element_text(hjust = 0),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.caption = element_text(hjust = 0, face= "italic")

     ) +
   NULL

patchwork::wrap_plots(plot_cmlinc_pw, plot_riskdiff_pw, plot_hazard_pw, ncol=1)
```

\newpage

### Figure S3: Second dose interval

Second dose events are censored by study end date, de-registration, death, and additionally by outcome-specific events.

```{r seconddose, echo=FALSE, warning=FALSE, message=FALSE}
knitr::include_graphics(fs::path(output_dir, "descriptive", "seconddose", "plot_seconddose_all.png"))
```

\newpage

### Figure S4: Event rates by calendar time

```{r eventrates, echo=FALSE, warning=FALSE, message=FALSE}
knitr::include_graphics(fs::path(output_dir, "descriptive", "eventdate", "plot_eventdate_step.png"))
```

\newpage

### Table S1: Inclusion criteria

```{r flowchart, echo=FALSE, warning=FALSE, message=FALSE}
flowchart %>%
  transmute(
    Participants = str_wrap(criteria, 40),
    N = n,
    `N excluded` = n_exclude,
    #`% excluded` = pct_exclude,
    `% remaining` = pct_all
  ) %>%
  mutate(
    Participants=if_else(row_number()==1,"HCWs aged 18-64 receiving BNT162b2 or ChAdOx1 between 4 Jan and 28 Feb",  Participants)
  ) %>%
  gt() %>%
  fmt_percent(
    columns = starts_with(c("%")),
    decimals = 1
  ) %>%
  fmt_number(
    columns = starts_with(c("N")),
    decimals= 0
  ) %>%
  fmt_missing(
    everything(),
    missing_text="-"
  ) %>%
  tab_style(
    style=cell_text(whitespace ="pre-wrap"),
    #style=cell_text(style ="italic"),
    locations = cells_body(columns = "Participants")
  )
```

\newpage

### Table S2: Baseline characteristics before eligibility exclusions

Characteristics for all HCWs aged 18-64 receiving a first dose of BNT162b2 or ChAdOx1 between 4 January and 28 February 2021 and actively registered at at TPP practice are reported, before exclusions due to missing demographic characteristics and Clinically Extremely Vulnerable status.

```{r table1_allvax, echo=FALSE, warning=FALSE, message=FALSE, out.extra=''}
table1_body <- read_csv(fs::path(output_dir, "descriptive", "tables", "table1_allvax.csv"))
table1_by <- read_csv(fs::path(output_dir, "descriptive", "tables", "table1_allvax_by.csv"))

table1 <- 
  table1_body %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(table1_by$by_col, table1_by$by_chr))

table1 %>%
  gt() %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
     label=md("Characteristic")
  ) %>%
  cols_hide(c("variable", "var_label", "order"))
```

\newpage

### Table S3: Event rates

```{r table_irr, echo=FALSE, message=FALSE, warning=FALSE}
table_irr <- read_csv(here("output", "descriptive", "tables", "table_irr.csv"))

table_irr %>%
  filter(
    event %in% c(
      "postest",
      "covidemergency",
      "covidadmitted",
      "coviddeath",
      "seconddose"
    )
  ) %>%
  transmute(
    outcome_descr, timesincevax,
    pfizer_q,
    pfizer_rate_fmt = if_else(pfizer_rate<0.001, "<0.001", number_format(0.001)(pfizer_rate)),
    pfizer_rate_fmt = if_else(pfizer_rate==0, "0", pfizer_rate_fmt),
    az_q,
    az_rate_fmt = if_else(az_rate<0.001, "<0.001", number_format(0.001)(az_rate)),
    az_rate_fmt = if_else(az_rate==0, "0", az_rate_fmt),
  ) %>%
  gt(
    groupname_col = "outcome_descr",
  ) %>%
  cols_label(
    outcome_descr = "Outcome",
    timesincevax = "Days since first dose",

    pfizer_q = "Events / person-years",
    pfizer_rate_fmt = "Incidence",

    az_q   = "Events / person-years",
    az_rate_fmt = "Incidence",
  ) %>%
  fmt_missing(
    everything(),
    missing_text="--"
  ) %>%
  cols_align(
    align = "right",
    columns = everything()
  ) %>%
  cols_align(
    align = "left",
    columns = "timesincevax"
  ) %>% 
  tab_spanner(
    label = "BNT162b2",
    columns = starts_with("pfizer")
  ) %>%
  tab_spanner(
    label = "ChAdOx1",
    columns = starts_with("az")
  )
```

### Table S4: Vaccine type for second doses

```{r seconddose_table, echo=FALSE, message=FALSE, warning=FALSE}
seconddose <- read_csv(fs::path(output_dir, "descriptive", "seconddose", "seconddose.csv"))
seconddose %>%
  select(-REDACTED, -pct_nonmiss) %>%
  gt(
    #groupname_col="First dose"
  ) %>%
  fmt_percent(
    columns = starts_with("pct"),
    decimals = 1
  ) %>%
  fmt_number(
    columns = starts_with(c("n")),
    decimals= 0
  ) %>%
  fmt_missing(
    everything(),
    missing_text="(REDACTED)"
  ) %>%
  cols_label(
    n = "N",
    pct = "%"
  )
```